### 2021-06-04 chapters/4_mutation_tool.tex
Starting with rq1 writeup
WIP exp1 design + finished characteristics

\chapter{Mutation testing tools}
\label{ch:mutation-tool}

@@

At the time of writing this thesis the studies are at least four years old. 

@@

Tools sometimes implement the same mutator differently, resulting in different mutant sets. Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT. This overview takes into account the mutators that are not already in the studies of Kintis et al., and Marki and Lindstrom. However some of the new mutators map to the existing mutators.

\begin{table}[htb]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{ MuJava mutator}}      & \multicolumn{2}{c|}{\textbf{PIT mutator}}                         \\ \hline
AORB & Arithmetic Operator Replacement Binary      & \multirow{4}{*}{M}   & \multirow{4}{*}{Math}                      \\ \cline{1-2}
ASRS & Short-Cut Assignment   Operator Replacement &                      &                                            \\ \cline{1-2}
SOR  & Shift Operator   Replacement                &                      &                                            \\ \cline{1-2}
COR  & Conditional Operator   Replacement          &                      &                                            \\ \hline
AOIU & Arithmetic Operator   Insertion Unary       & \multirow{2}{*}{UOI} & \multirow{2}{*}{Unary Operation Insertion} \\ \cline{1-2}
AOIS & Arithmetic Operator   Insertion Short-cut   &                      &                                            \\ \hline
ODL  & Operator Deletion                           & AOD                  & Arithmic Operator   Deletion               \\ \hline
AODS & Arithmetic Operator   Deletion Short-cut    & RI                   & Remove Increments                          \\ \hline
ROR  & Relational Operator   Replacement           & CB                   & Conditionals Boundary                      \\ \hline
AORS & Arithmetic Operator   Replacement Short-Cut & I                    & Increments                                 \\ \hline
COD  & Conditional Operator   Deletion             & RC                   & Remove Conditionials                       \\ \hline
\end{tabular}
\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJavav and PIT}
\end{table}
Next we tried to generate mutants with all three tools. 
All of the sources that we used for generating mutants are build with the build tools Maven or Gradle. 
Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major includes support for Maven, but it is not documented anywhere.
For this reason we could not generate mutants with Major.
Table \ref{tab:mutants-generated} shows the amount of mutants generated per project and per tool. The results show that in all of the cases PIT generates more mutants than MuJava. In some cases the difference is significant. This difference can partially be explained by the fact that MuJava does not support source projects with Java version 1.7 or higher\cite{mujava}. All the the source code that uses features from this version of Java and higher resulted in an error during mutant generation.

\begin{table}[htb]
\centering
\begin{tabular}{|l|l|c|c|c|}
\hline
                                & \textbf{version} & \textbf{\begin{tabular}[c]{@{}c@{}}LOC\\ (Lines of code)\end{tabular}} & \multicolumn{1}{l|}{\textbf{MuJava}} & \multicolumn{1}{l|}{\textbf{PIT}} \\ \hline
org.apache.commons.numbers.core & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
Jodatime                        & 2.10.10          & 28811                                                                  & 52925                                & 112772                            \\ \hline
Zxing 3.4.1                     & 3.4.1            & 24792                                                                  & 65983                                & 161409                            \\ \hline
Google Auto Common              & 0.11             & 2338                                                                   & 20                                   & 5219                              \\ \hline
Google Auto Factory             & 1.0-beta9        & 1507                                                                   & 69                                   & 5832                              \\ \hline
Google Auto Value               & 1.7.5            & 7466                                                                   & 745                                  & 16746                             \\ \hline
Google Auto Service             & 1.0-rc6          & 750                                                                    & 0                                    & 475                               \\ \hline
ScribeJava-Core                 & 8.1.0            & 5709                                                                   & 1358                                 & 5746                              \\ \hline
Checkstyle                      & 8.41.1           & 38491                                                                  & 4529                                 & 100952                            \\ \hline
fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
\end{tabular}
\caption{\label{tab:mutants-generated}Amount of mutants generated by MuJava and PIT}
\end{table}
\newpage
\section{Mutation tool}
We have studied literature on three different tools and extended the research with more recent data.
MuJava is not actively maintained and has not been updated in the last few years.
It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
These are test engines\todo{add to vocabulary} and are crucial for executing tests.
MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
Conforming to these requirements would recude the set of projects we could use in or experiments.
\newline
While Major supports JUnit 4,
we did not succeed in generating mutants with this tool.
It would be too time consuming to customize all source projects to work with Major.
Furthermore there is not much documentation available for this tool.
\newline
PIT targets the industry, is actively maintained and is open source\cite{Kintis2016AnalysingStudy}.
It supports Maven, Gradle, has a \acrfull{cli} and has a faster execution time than the other tools.
For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing\cite{pit}.
PIT also generated significantly more mutants in every project we have tested.
Based on the information presented in this chapter we dcecided that PIT would be the best choice for generating and executing mutants.
\end{document}

>>>>

\chapter{Mutation testing tools}
\label{ch:mutation∂-tool}

@@

At the time of writing this thesis these studies are at least four years old. 

@@

Tools sometimes implement the same mutator differently, resulting in different mutant sets. Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT. This overview does include the mutators that have been reviewed in the studies of Kintis et al., and Marki and Lindstrom. However some of the new mutators map to the existing mutators.

\begin{table}[htb]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{ MuJava mutator}}      & \multicolumn{2}{c|}{\textbf{PIT mutator}}                         \\ \hline
AORB & Arithmetic Operator Replacement Binary      & \multirow{4}{*}{M}   & \multirow{4}{*}{Math}                      \\ \cline{1-2}
ASRS & Short-Cut Assignment   Operator Replacement &                      &                                            \\ \cline{1-2}
SOR  & Shift Operator   Replacement                &                      &                                            \\ \cline{1-2}
COR  & Conditional Operator   Replacement          &                      &                                            \\ \hline
AOIU & Arithmetic Operator   Insertion Unary       & \multirow{2}{*}{UOI} & \multirow{2}{*}{Unary Operation Insertion} \\ \cline{1-2}
AOIS & Arithmetic Operator   Insertion Short-cut   &                      &                                            \\ \hline
ODL  & Operator Deletion                           & AOD                  & Arithmic Operator   Deletion               \\ \hline
AODS & Arithmetic Operator   Deletion Short-cut    & RI                   & Remove Increments                          \\ \hline
ROR  & Relational Operator   Replacement           & CB                   & Conditionals Boundary                      \\ \hline
AORS & Arithmetic Operator   Replacement Short-Cut & I                    & Increments                                 \\ \hline
COD  & Conditional Operator   Deletion             & RC                   & Remove Conditionials                       \\ \hline

\end{tabular}
\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJavav and PIT}
\end{table}
Next we tried to generate mutants with all three tools. 
All of the sources that we used for generating mutants are build with the build tools Maven or Gradle. 
Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major includes support for Maven, but it is not documented anywhere.
For this reason we could not generate mutants with Major.
Table \ref{tab:mutants-generated} shows the amount of mutants generated per project and per tool. The results show that in all of the cases PIT generates more mutants than MuJava. In some cases the difference is significant. This difference can partially be explained by the fact that MuJava does not support source projects with Java version 1.7 or higher\cite{mujava}. All the the source code that uses features from this version of Java and higher resulted in an error during mutant generation.
\begin{tabular}{|l|l|c|c|c|}
                                & \textbf{version} & \textbf{\begin{tabular}[c]{@{}c@{}}LOC\\ (Lines of code)\end{tabular}} & \multicolumn{1}{l|}{\textbf{MuJava}} & \multicolumn{1}{l|}{\textbf{PIT}} \\ \hline
Commons-numbers.core            & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
Jodatime                        & 2.10.10          & 28811                                                                  & 52925                                & 112772                            \\ \hline
Zxing 3.4.1                     & 3.4.1            & 24792                                                                  & 65983                                & 161409                            \\ \hline
Google Auto Common              & 0.11.0           & 2338                                                                   & 20                                   & 5219                              \\ \hline
Google Auto Factory             & 1.0-beta9        & 1507                                                                   & 69                                   & 5832                              \\ \hline
Google Auto Value               & 1.7.5            & 7466                                                                   & 745                                  & 16746                             \\ \hline
Google Auto Service             & 1.0-rc6          & 750                                                                    & 0                                    & 475                               \\ \hline
ScribeJava-Core                 & 8.1.0            & 5709                                                                   & 1358                                 & 5746                              \\ \hline
Checkstyle                      & 8.41.1           & 38491                                                                  & 4529                                 & 100952                            \\ \hline
Fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
Jfreechart                      & 1.5.3            & 91876                                                                  & 0                                    & 350741                            \\ \hline
Commons-lang                    & 3.12.0           & 29836                                                                  & 31145                                & 134764                            \\ \hline
Commons-codec                   & 1.15.0           & 9656                                                                   & 21719                                & 54804                            \\ \hline
Commons-text                   & 1.9.0             & 9781                                                                   & 10403                                & 48490                            \\ \hline
Commons-io                     & 2.8.0             & 13947                                                                  & 13245                                & 44631                            \\ \hline
Gson                           & 2.8.6             & 8078                                                                   & 9198                                 & 28485                            \\ \hline
Commons-cli                    & 1.4.0             & 2782                                                                   & 831                                  & 7193                             \\ \hline
Commons-csv                    & 1.8.0             & 1855                                                                   & 2081                                 & 6906                             \\ \hline
\caption{\label{tab:mutants-generated}Amount of mutants generated by MuJava and PIT}
\section{Mutation tool}
We have studied literature on three different tools and extended the research with more recent data.
MuJava is not actively maintained and has not been updated in the last few years.
It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
These are test engines\todo{add to vocabulary} and are crucial for executing tests.
MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
Conforming to these requirements would recude the set of projects we could use in or experiments.
While Major supports JUnit 4,
we did not succeed in generating mutants with this tool.
It would be too time consuming to customize all source projects to work with Major.
Furthermore there is not much documentation available for this tool.
PIT targets the industry, is actively maintained and is open source\cite{Kintis2016AnalysingStudy}.
It supports Maven, Gradle, has a \acrfull{cli} and has a faster execution time than the other tools.
For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing\cite{pit}.
PIT also generated significantly more mutants in every project we have tested.
Based on the information presented in this chapter we dcecided that PIT would be the best choice for generating and executing mutants.
\end{document}

### 2021-06-04 chapters/5_identifying_characteristics.tex
WIP exp1 design + finished characteristics

\chapter{Identifying characteristics}
\label{ch:clustering_characteristics}
Our goal is to cluster mutants based on characteristics. 
To achieve this we need to identify characteristics that the mutants contain. 
In this chapter we identify characteristics which we can use for clustering.

\section{Overview of characteristics}
Zhang et al., identifies several mutant characteristics\cite{Zhang2019PredictiveTesting}. 
Their research focuses on detecting if a mutant survives or not before executing the mutant.
While the goal of their research is different than ours, we can still use the characteristics they identified.
We select characteristics relevant to our research from the research of Zhang et al.
We extend the list of characteristics with data we can extract from our selected mutation testing tool. 
\newline
PIT generates and executes mutants on byte code level\cite{pitestBytecode}.
By mutating in Java byte code we can identify characteristics specific to Java byte code.
PIT also gathers mutant metadata.
We can extract this mutant metadata and identify characteristics. 
With the combination of information extracted from PIT and characteristics selected from existing research we identify the following characteristics:
\begin{enumerate}
    \item Mutant similarity.
    \item Amount of tests the mutant is challenged by.
    \item Mutator identifier.
    \item Mutant opcode.
    \item Mutant return type.
    \item Mutant contains local variables.
    \item Mutant is in try catch block.
    \item Mutant is in finally block.
    \item Number of mutated byte code instructions.
    \item Name of class that contains the mutant.
    \item Name of function that contains the mutant.
    \item Line number of block that contains the mutant.
    \item Line number on which the mutation occur.
    \item Number of lines of byte code of mutated class.
\end{enumerate}
In the following subsections we will elaborate more on these characteristics.

\section{Mutant similarity}
A piece of code is a textual representation of instructions for a computer.
A mutant is a transformation of a piece of code.
This transformation is predefined.
The difference between a mutant and its parent can be defined as similarity.
We identify the similarity between a mutant and parent as a characteristic.
\newline
To measure this characteristic we select a similarity measure.
There are existing similarities measures that base their similarity in different ways\cite{Gomma2013AApproaches}. 
We select the Levenshtein distance as similarity measure for our research\cite{Gomma2013AApproaches}.
The Levenshtein distance defines the distance between two strings by counting the minimum number of operations(create/edit/delete) needed to transform one string into the other\cite{Gomma2013AApproaches}.
The more the textual representation of a mutant differs from its parent the bigger the Levenshtein distance will be.

\section{Amount of tests challenged by mutant}
A mutant may be challenged by tests.
If the test fails, the mutant is considered killed.
If the test passes, the mutant is considered survived.
If the mutant is not challenged by any tests it is also considered survived.
The amount of tests a mutant is challenged by is dependant on the coverage and test effectiveness of the test suite.
We identify this number as a characteristic.

\section{Mutator identifier}
At the time of writing the selected mutation testing tool PIT supports more than 100 mutators\cite{pitestMutators}.
A mutator is the definition of how the code is mutated from its original\cite{Offutt2001}.
Each mutator mutates the source code in a dfferent way.
PIT assigns identifiers to their mutators for execution and reporting purposes. 
Each mutant has a mutator identifier that matches with how the mutant is mutated from the source code.
We can extract this data from PIT and use it as a characteristic.

\section{Mutant opcode}
\say{A Java Virtual Machine instruction consists of a one-byte opcode specifying the operation to be performed, followed by zero or more operands supplying arguments or data that are used by the operation.} (Java 12 SE docs ch.12).
When PIT mutates the source code it uses Java opcodes to alter the original byte code. 
A record of which opcode is being used is stored in memory during the mutation testing process. 
This opcode can differ per mutator and is dependant on how the code is written.
During the mutation testing process of PIT we can extract the opcode.
We identify the opcode as a characteristic.

\section{Mutant return type}
A mutant may contain a return statement. 
This is dependant on the source code that is mutated.
A return statement can be identified as a characteristic.
If a mutant contains a return statement, we can extract its type from PIT.
There is the possibility that a mutant does not contain a return statement, in this case we will set the mutant return type to a default value.

\section{Mutant location}
The location of a mutant in PIT is defined by multiple properties;
the name of the class, the name of the function, the line number of the start of the block the mutant is located in and the line number the mutation occurs on.
These properties refer to the location in Java byte code.
The location can be identified as a characteristic of a mutant.
PIT uses this location in the mutation testing and reporting processes.
We can extract the location as defined by the properties in PIT.
Each property is identified as a characters

\section{Extracting characteristics}
PIT exposes some of the characteristics with its plugin system.
Unfortunately not all characteristics are readily available for extraction.
To extract all the characteristics from PIT we create a modified version of the application.
We adjust the existing code to export the characteristics per mutant in various stages of the mutation testing process.
We create a new model that stores the listed characteristics trough out the mutation testing process. 
The data can be written to a file during the reporting stage of the process.

>>>>

\chapter{Identifying characteristics}
\label{ch:clustering_characteristics}
Our goal is to cluster mutants based on characteristics. 
To achieve this we need to identify characteristics that the mutants contain. 
In this chapter we identify characteristics that are relevant for defining clusters.

\section{Overview of characteristics}
Zhang et al., identifies several mutant characteristics\cite{Zhang2019PredictiveTesting}. 
Their research focuses on detecting if a mutant survives or not before executing the mutant.
The mutant characteristics identified by Zhang et al., have been used in other research with promising results\cite{Oonk2021}.
While the goal of their research is different than ours, we can still use the characteristics they identified.
We select characteristics relevant to our research from the research of Zhang et al and Oonk.
We extend the list of characteristics with data we can extract from our selected mutation testing tool. 
\newline
PIT generates and executes mutants on byte code level\cite{pitestBytecode}.
By mutating in Java byte code we can identify characteristics specific to Java byte code.
PIT also gathers mutant metadata.
We can extract this mutant metadata and identify characteristics. 
With the combination of information extracted from PIT and characteristics selected from existing research we identify the following characteristics:
\begin{enumerate}
    \item Mutant similarity.
    \item Amount of tests the mutant is challenged by.
    \item Mutator identifier.
    \item Mutant opcode.
    \item Mutant return type.
    \item Mutant contains local variables.
    \item Mutant is in try catch block.
    \item Mutant is in finally block.
    \item Name of class that contains the mutant.
    \item Name of function that contains the mutant.
    \item Line number of the start of the block that contains the mutant.
    \item Line number on which the mutation occur.
\end{enumerate}
In the following subsections we will elaborate more on these characteristics.

\section{Mutant similarity}
A piece of code is a textual representation of instructions for a computer.
A mutant is a transformation of a piece of code.
This transformation is predefined.
The difference between a mutant and its parent can be defined as similarity.
We identify the similarity between a mutant and parent as a characteristic.
\newline
To measure this characteristic we select a similarity measure.
There are existing similarities measures that base their similarity in different ways\cite{Gomma2013AApproaches}. 
We select the Levenshtein distance as similarity measure for our research\cite{Gomma2013AApproaches}.
The Levenshtein distance defines the distance between two strings by counting the minimum number of operations(create/edit/delete) needed to transform one string into the other\cite{Gomma2013AApproaches}.
The more the textual representation of a mutant differs from its parent the bigger the Levenshtein distance will be.
A unit test tests a small piece of code.
If this piece of codes differs too much from what is being tested the test is more likely fail. 
The Levenshtein distance measures this difference.
This could be a useful characteristic because mutants that have the same similarity could be grouped together.

\section{Amount of tests challenged by mutant}
A mutant may be challenged by tests.
If the test fails, the mutant is considered killed.
If the test passes, the mutant is considered survived.
If the mutant is not challenged by any tests it is also considered survived.
The amount of tests a mutant is challenged by is dependant on the coverage and test effectiveness of the test suite.
We identify this number as a characteristic.
Every test the mutant is challenged by is an occasion for the mutant to be killed or not.
This could be a useful characteristic because mutants that have more chances to be killed could be grouped together.

\section{Mutator identifier}
At the time of writing the selected mutation testing tool PIT supports more than 100 mutators\cite{pitestMutators}.
A mutator is the definition of how the code is mutated from its original\cite{Offutt2001}.
Each mutator mutates the source code in a dfferent way.
PIT assigns identifiers to their mutators for execution and reporting purposes. 
Each mutant has a mutator identifier that matches with how the mutant is mutated from the source code.
We can extract this data from PIT and use it as a characteristic.
It could be a useful characteristic because mutants that are changed in the same predefined way could be grouped together.

\section{Mutant opcode}
\say{A Java Virtual Machine instruction consists of a one-byte opcode specifying the operation to be performed, followed by zero or more operands supplying arguments or data that are used by the operation.} (Java 12 SE docs ch.12).
When PIT mutates the source code it uses Java opcodes to alter the original byte code. 
A record of which opcode is being used is stored in memory during the mutation testing process. 
This opcode can differ per mutator and is dependant on how the code is written.
During the mutation testing process of PIT we can extract the opcode.
We identify the opcode as a characteristic.
The opcode could be a useful characteristic because mutants with the same opcode may be grouped together.

\section{Mutant return type}
A mutant may contain a return statement. 
This is dependant on the source code that is mutated.
A return statement can be identified as a characteristic.
If a mutant contains a return statement, we can extract its type from PIT.
There is the possibility that a mutant does not contain a return statement, in this case we will set the mutant return type to a default value.
The return type is a distinctive property of a piece of code, this distinction a useful characteristic because mutants with the same return type may be grouped together.

\section{Mutant location}\label{ch:characteristics_location}
The location of a mutant in PIT is defined by multiple properties;
the name of the class, the name of the function, the line number of the start of the block the mutant is located in and the line number the mutation occurs on.
These properties refer to the location in Java byte code.
The location can be identified as a characteristic of a mutant.
PIT uses this location in the mutation testing and reporting processes.
We can extract the location as defined by the properties in PIT.
Each property is identified as a characters
The location of a piece of code is a distinctive property. 
The same piece of code in a different location may have a different effect on the source code or unit tests.
A mutant may not be unique if the location is not taken into account.
It could be a useful characteristic because mutants that are in the same location may be grouped together.

\section{Error handling capability}
A technique to handle errors in Java is to use try-catch blocks\cite{javaDocs}. 
The presence of try-catch blocks are an indication of the error handling capabilities of certain pieces of code.
An addition to the try-catch blocks are finally blocks.
A finally block is executed after a try and/or catch block\cite{javaDocs}.
There is a possibility that a mutant may be generated inside these blocks.
If a mutant is located in a try, catch or finally block can be defined as a characteristic.
It could be a useful characteristic because the mutants that are located in the same type of block may be grouped together.

\section{Local variables}
A mutant may be located in a location(see Chapter \ref{ch:characteristics_location}) that may contain local variables.
There is a chance a mutant modifies a local variable.
A variable modified by a mutant increases the chance for a mutant to be killed.
This is because a unit test may check for a specific value for a specific local variable.
The amount of local variables a location contains can be identified as a characteristic.
We can extract the amount of local variables a location contains from PIT.
This could be a useful characteristic as mutants with the same amount of local variables may be grouped together.

### 2021-06-04 chapters/6_experiment_design.tex
WIP exp1 design + finished characteristics

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Experiment design}
\todo{expand this}
We hypothesise that there is a we can cluster on the characteristics of a mutant such that each mutant should be a representative for all of the mutants in its cluster.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.

We design an experiment in which we implement the following steps:
\begin{enumerate}
    \item Extract all the characteristics.
    \item For each mutant calculate the Levenshtein distance.
    \item Cluster the mutants based on the characteristics.
    \item Randomly select one mutant from each cluster.
    \item Execute the selected mutants in PIT and calculate the mutation score.
\end{enumerate}
We elaborate on these steps in the following subsections.

\subsection{Levenshtein distance on Java byte code}
A mutant is a piece of code that differs in a predefined way from its parent.
Java source code compiles into Java byte code.
During the compilation of Java code certain context is abstracted away\cite{byteCodeEngineering}.
Optimisations are applied which also changes the Java byte code\cite{byteCodeEngineering}.
As a result the textual similarity between a Java code mutant and its parent is different from the similarity between a Java byte code mutant and its parent.
While the textual similarity may be different between the Java code and Java byte code the functionality remains the same.
Java byte code reflects more of the semantic nature of the source code than the source code itself does. 
\newline
PIT generates and executes mutants on byte code level\cite{pitestBytecode}. 
In other words PIT executes the unit tests of a source against the byte code of a mutant.
\newline
For this experiment we will use the Levenshtein distance between a Java byte code mutant and its parent.
By calculating the distance on byte code we filter out the context that is present in Java code.
This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.

@@

\subsection{Hierarchical clustering}
Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
Hierarchical clustering is subdivided into agglomerative and divisive. 
The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011HierarchicalStudy}. 
The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011HierarchicalStudy}.
\newline
Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
\newline
We identify each Levenshtein distance between a mutant and its parent as a separate cluster.
Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented as Levenshtein distances.
\todo{write about implementation}

\subsection{Weighed mutation score}
Our hypothesis states that each mutants should be a representative for all the mutants in its cluster. 
This translates to if a mutant survives or is killed all mutants in that cluster should survive or be killed respectively.
To reflect this in our experiment a weighed mutation score is calculated for the clustered mutants.
This weighed mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
The weighed mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
\newline
For example, take a full set with a score of 75/100 killed mutants. \todo{example too much?}
This gives us a mutation score of 75\%. 
We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
We randomly select four mutants of each cluster and execute them.
The mutants representing cluster one and four survive.
If we calculate the weighed score we get 68/100 which is 68\%.
We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).

\subsection{PIT plugin system}
PIT offers a plugin system in which developers can inject their own functionality in PIT\cite{pitestPlugin}. 
There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
A mutation result listener  receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
A Mutation interceptor passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
The implementation of our experiment makes use of this plugins system.
\newline
We developed three separate plugins for our experiment.
The first plugin is a Mutation interceptor.
It analyses each mutant, calculates the Levenshtein distance and writes a mutant identifier and distance to a \acrshort{csv} file. 
It also clears the list of mutants provided by PIT to prevent execution. 
The second plugin is also a Mutation interceptor.
It reads the clustered mutants from a \acrshort{csv} file and compares them to the list of mutants provided by PIT.
Only the clustered mutants are passed back to PIT for execution.
The last plugin is a mutation result listener. 
This plugin calculates the weighed mutation score and outputs it to a text file.
\newline
The plugins in PIT are executed based on type and per source class.
This means that a plugin will never receive all the mutants generated for a whole project during execution.
Because of how PIT executes its plugins we have to execute PIT twice.
In the first run we calculate the Levenshtein distance for all mutants and in the second run we execute the selected mutants and calculate the weighed mutation score.
\end{document}

>>>>

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Experiment design rq1}\todo{discuss how to split this with exp design for rq2}
In this chapter we will design an experiment to answer our first research question.
We hypothesise that we can cluster mutants with the identified set of characteristics 
while maintaining effectiveness when executing one mutant from each cluster that is in the majority of that cluster.
If a cluster contains ten mutants with eight mutants killed and two mutants survived, all the mutants that are killed are in the majority of that cluster.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
\newline
We will start with extracting the characteristics from the selected mutation testing tool. 
Next we select and implement a clustering algorithm for which we can cluster mutants.
We can then cluster the mutants and compare the results of the clusters with the results of the mutants.

\section{Extracting characteristics}
To extract all the characteristics we identified we need to run PIT and configure it to generate all mutants that are possible within the tool. 
PIT works in phases, the first phase is the generation phase where mutants are generated.
The second phase is the execution phase where mutants are challenged by the test suite.
The last phase is the reporting phase where a report is generated based on the users preferences and data generated by PIT.
During the various stages we can extract characteristics by using PIT's plugin system.
\newline
PIT offers a plugin system in which developers can inject their own code in PIT\cite{pitestPlugin}.
There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
A mutation result listener receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
A Mutation interceptor is passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
The implementation of our experiment makes use of this plugins system.
\newline
We developed two plugins to extract characteristics.
The first plugin is of type Mutation interceptor.
It is provided a list of details per class. 
We can extract all characteristics but the Levensthein distance and the amount of tests a mutant is challenged by.
To calculate the Levensthein distance we need the byte code of the original code and the mutated code.
We generate the byte code of the original code based on the location provided in the list of details.
We make use of existing functionality in PIT for generating byte code.
The mutator is also provided in the list of details.
With the combination of location and mutator we can generate the byte code of the mutant.
With both pieces of code available we can calculate the Levenshtein distance between them.
\newline
The characteristics gathered until this phase of the mutation testing process are written, together with a unique identifier, to a \acrshort{csv} file.
\newline
The second plugin is a mutation result listener. 
This plugin gets passed a list of the results per mutant.
This list contains data about the mutant survival status.
It also contains data about the number of tests the mutant is challenged by.
We store this data together with the identifier to a different \acrshort{csv} file.
\newline
When the mutation testing process has finished we need to merge the characteristic gathered in the second plugin.
We add the characteristic; the number of tests the mutant is challenged by to the data file containing all the other characteristics.
We can merge the data because both files contain the same unique identifier per mutant.

\section{Levenshtein distance on Java byte code}
A mutant is a piece of code that differs in a predefined way from its parent.
Java source code compiles into Java byte code.
During the compilation of Java code certain context is abstracted away\cite{byteCodeEngineering}.
Optimisations are applied which also changes the Java byte code\cite{byteCodeEngineering}.
As a result the textual similarity between a Java code mutant and its parent is different from the similarity between a Java byte code mutant and its parent.
While the textual similarity may be different between the Java code and Java byte code the functionality remains the same.
Java byte code reflects more of the semantic nature of the source code than the source code itself does. 
PIT generates and executes mutants on byte code level\cite{pitestBytecode}. 
In other words PIT executes the unit tests of a source against the byte code of a mutant.
\newline
For our characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
By calculating the distance on byte code we filter out the context that is present in Java code.
This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.

@@

\section{Hierarchical clustering}\todo{elaborate on why hierarchical clustering}
Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
Hierarchical clustering is subdivided into agglomerative and divisive. 
The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011HierarchicalStudy}. 
The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011HierarchicalStudy}.
\newline
Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
\newline
We identify all characteristics per mutant as a separate cluster.
Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented as Levenshtein distances.
\todo{write about implementation}

\section{Weighed mutation score}
Our hypothesis states that each mutants executed should be in the majority of mutants in its cluster. 
With a mutant executed from each cluster we can calculate a mutation score.
This mutation score is a weighed mutation score.
This weighed mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
The weighed mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
\newline
For example, take a full set with a score of 75/100 killed mutants. \todo{example too much?}
This gives us a mutation score of 75\%. 
We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
We randomly select four mutants of each cluster and execute them.
The mutants representing cluster one and four survive.
If we calculate the weighed score we get 68/100 which is 68\%.
We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).

\end{document}
